Query Optimization for Dynamic Imputation

### 2017-03-01 paper/experiments.tex
paper: removed table names from missing val tables
Incorporate feedback on Imputation Strategies section

The original tables have a large number of attributes, in some cases providing more granular tests results or alternative metrics.
We focused on a subset of the attributes for each table to simplify the presentation and exploration of queries.
\Cref{table:nhanes-description} shows the attributes selected, along with the percentage of \nullv{} values for each attribute.
For readability, we have replaced the NHANES variable names with self-explanatory attribute names.

\begin{table}
  \centering
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_demo}
    \caption{Demographics. \demorows{} rows.}
  \end{subtable}
  \par\medskip
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_labs}
    \caption{Laboratory Results. \labexrows{} rows.}
  \end{subtable}
  \par\medskip  
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_exams}
    \caption{Physical Results. \labexrows{} rows.}
  \end{subtable}
  \par\medskip  
  \caption{Missing value distribution for each table/attribute in CDC NHANES 2013--2014 data.}\label{table:nhanes-description} 
\end{table}

@@

\begin{table}
  \centering
  \input{tables/fcc}
  \caption{Missing value distribution for each attribute in freeCodeCamp Survey Data}\label{table:fcc-description} 
\end{table}

@@

%\begin{figure}
%\includegraphics[width=\columnwidth]{figures/planning_times_imputedb.png}
%\caption{Planning times for each query}
%\label{fig:plantimes}
%\end{figure}

\subsubsection{Accuracy vs. Base-table Imputation}

>>>>

The original tables have a large number of attributes, in some cases providing more granular tests results or alternative metrics.
We focused on a subset of the attributes for each table to simplify the presentation and exploration of queries.
\Cref{table:nhanes-description} shows the attributes selected, along with the percentage of \nullv{} values for each attribute.
For readability, we have replaced the NHANES attribute names with self-explanatory attribute names.

\begin{table}
  \centering
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_demo}
    \caption{Demographics (\texttt{demo}). \demorows{} rows.}
  \end{subtable}
  \par\medskip
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_labs}
    \caption{Laboratory Results (\texttt{labs}). \labexrows{} rows.}
  \end{subtable}
  \par\medskip  
  \begin{subtable}{0.5\textwidth}
    \centering
    \input{tables/cdc_exams}
    \caption{Physical Results (\texttt{exams}). \labexrows{} rows.}
  \end{subtable}
  \par\medskip  
  \caption{Missing value distribution for each table/attribute in CDC NHANES 2013--2014 data.}\label{table:nhanes-description} 
\end{table}

@@

\begin{table}
  \centering
  \input{tables/fcc}
  \caption{Missing value distribution for each attribute in freeCodeCamp Survey Data (\texttt{fcc}}\label{table:fcc-description} 
\end{table}

@@

%\begin{figure}
%\includegraphics[width=\columnwidth]{figures/planning_times_imputedb.png}
%\caption{Planning times for each query}
%\label{fig:plantimes}
%\end{figure}
\vspace{0.5in}
\subsubsection{Accuracy vs. Base-table Imputation}

### 2017-03-01 paper/main.tex

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for the usability of data. Users
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by applying a potentially-expensive imputation strategy. Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, 
performing any operations needed on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations. This placement is based
on the specific query issued and a user-provided parameter that trades off data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}. Our
results show that the optimizer generates query plans that execute between \lowxalphazero{}
and \highxalphazeroexacs{} times as fast as imputing at the base tables and then executing
queries, making exploratory data analysis using our system feasible.  Furthermore, we show
that the query results from on-the-fly imputation have
error between \lowsmapealphazero{} and
\highsmapealphaoneexacs{} percent when compared to standard full-table imputation
approach.

\end{abstract}

>>>>

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for the usability of data. Users
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by applying a potentially-expensive imputation strategy. Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, 
performing any operations needed on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations. This placement is based
on the specific query issued and a user-provided parameter that trades off data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}. Our
results show that the optimizer generates query plans that execute between \lowxalphazero{}
and \highxalphazeroexacs{} times as fast as imputing at the base tables and then executing
queries, making exploratory data analysis using our system feasible.  Furthermore, we show
that the query results from on-the-fly imputation have
error between \lowsmapealphazero{} and
\highsmapealphaoneexacs{} percent when compared to standard full-table imputation
approach. 
We calculate the number of tuples used to produce each aggregate
in our experiments, and find that when optimizing for quality between 0 to 10\% fewer tuples satisfy query constraints relative to the full-table imputation. When all tuples with missing values are dropped, the aggregate
is computed with 3 to 88\% fewer tuples.
\end{abstract}
